Circumnuclear Media of Quiescent Supermassive Black Holes

### 2015-07-02 ms.tex
sclaing relations and caveats

\section{Model}
\label{sec:model}
\subsection{Galaxy models}
\label{sec:gal_model}
\citet{LauerFaber+:2007a} use Hubble Space Telescope WFPC2 imaging to
measure the radial surface brightness profiles for hundreds of nearby
early type galaxies. The measured profile is fit by a ``Nuker" law parameterization:
\begin{equation}
  I(\xi)=I_b 2^{(\beta-\Gamma)/\alpha} \xi^{-\Gamma} (1+\xi^\alpha)^{-(\beta-\Gamma)/\alpha}, \,\,\,\xi\equiv\frac{r}{r_b},
\end{equation}
i.e., a broken power law that transitions from an inner power law
slope, $\Gamma$, to an outer power law slope, $\beta$, at a break
radius, $\rb$.  If the stellar population is spherically symmetric,
then this corresponds to a 3D stellar density $\rhostar \propto
r^{-1-\Gamma}$ for $r \ll \rb$ and $\rhostar\propto r^{-1-\beta}$ for
$r \gg \rb$.  We can write the deprojected stellar density
approximately (formally, this is the $\alpha \rightarrow \infty$
limit) as
\begin{align}
\rho_\star = 
\begin{cases}
\rho_\star|_{\rinf} \left(r/\rinf\right)^{-1-\Gamma} & r \leq r_b\\
\rho_\star|_{r_b} \left(r/r_b\right)^{-1-\beta} & r > r_b,
\label{eq:rhostar}
\end{cases}
\end{align}
where $\rho_\star|_{\rinf}$ is the stellar density at the radius of
the black hole sphere of influence\footnote{This approximate
  expression agrees surprisingly well with the mean empirical scaling
  relation for $r_{\rm inf}(M_\bullet)$ calculated in Appendix C of
  \citet{Stone&Metzger15}, although we note that this relation has
  significant scatter. Also, we note that the  scaling is somewhat
  different for cores and cusps. Fixing the power law slope,
  $\rinf\simeq 25 (8) \Mbheight^{0.6}$ for core (cusp)
  galaxies. Unless otherwise noted we will use the single relation in
  equation~\eqref{eq:rsoi}},
\begin{equation}
  \rinf \simeq G \Mbh/\sigma_{\bullet}^2 \approx 14 M_{\bullet,8}^{0.6}\,{\rm pc},
\label{eq:rsoi}
\end{equation}
where $M_{\bullet,8} \equiv M_{\bullet}/10^{8}M_{\odot}$ and the
second equality in (\ref{eq:rsoi}) employs the $\Mbh-\sigma$
relationship of \citet{McConnellMa+:2011a},
 \begin{align}
M_{\bullet} \simeq 2\times 10^{8}\left(\frac{\sigma_{\bullet}}{200\,{\rm
      km\,s^{-1}}}\right)^{5.1}M_{\odot}.
\label{eq:Msigma}
\end{align}
This may be of questionable validity for low mass black holes (e.g.,
\citealt{Greene+2010, Kormendy+2013}). Also, several of the black hole
masses used in \citealt{McConnellMa+:2011a} were underestimated
\citep{Kormendy+2013}.

@@

We fix $\beta = 2$ but find that our results are not overly sensitive
to the properties of the gas flow on radial scales $\gtrsim r_b$ for
values of $\beta\gsim 1.5$, which characterize most of the
\citet{LauerFaber+:2007a} sample\footnote{The median $\beta$ in the
  \citet{LauerFaber+:2007a} sample is 1.39 (1.64) for core (cusp)
  galaxies}.  The presence of the break radius $r_{b}$ is, however,
necessary to obtain a converged steady state for some regions of our
parameter space.  We consider solutions calculated for up to four
values of $\rb$: 50 pc, 100 pc, 200 pc, and 400 pc, motivated by the
range of break radii from the \citet{LauerFaber+:2007a}
sample.\footnote{For the core galaxies the break radius follows the
  scaling relationship $\rb\sim 106 \, (\Mbheight)^{0.39}$ pc, with
  scatter of approximately one dex.  Most the cusp galaxies have $\rb$
  between 100 and 1000 pc, and lack a clear trend with $\Mbh$. {\bf AG
    I redid the fitting and found $\rb\sim 90 \, (\Mbheight)^{0.5}$
    pc}} We neglect values of $r_b$ which would give unphysically
large bulges for a given $\Mbh$.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{ti.pdf}
  \caption{Minimum effective wind heating parameter required for
    thermal stability as a function of SMBH mass.  Black lines show
    $v_{\rm TI}$ (eq.~[\ref{eq:cooling3}]), the heating rate required
    for $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}} > 10$ in the
    high-heating limit when the stagnation radius lies interior to the
    influence radius, for different values of the mass loss parameter
    $\eta$ as marked.  Blue lines show the minimum heating parameter
    required to have $\zeta > \zeta_c = (r_{b}/r_{\rm
      inf})^{0.5(1-\Gamma)}$ (eq.~[\ref{eq:zetacrit}]), separately for
    cusp (solid) and core (dashed) galaxies. Based on the
    \citet{LauerFaber+:2007a} sample we take $\rinf=25 (8) \Mbh^{0.6}
    pc$ and $\rb=106 \Mbheight^{0.4}$ pc (100 pc) for cores (cusps)
    {\bf AG update rb fit}. For $\zeta < \zeta_c$ the stagnation
    radius moves from inside the influence radius, out to the stellar
    break radius $r_b$ (Appendix \ref{app:be}); in practice renders
    the flow susceptible to thermal runaway, even if $v_{w} > v_{\rm
      TI}$.  For cusp galaxies the $v_w > v_{\rm TI}$ is usually the
    more strigent criterion for stability, while for core galaxies the
    $\zeta > \zeta_c$ is usually more relevant.}
\label{fig:TI}
\end{figure}

@@

Figure \ref{fig:TI} shows $v_{\rm TI}(M_{\bullet})$ for different mass
input parameters $\eta$, as well as the [$\eta$-independent] $\zeta >
\zeta_c$ criterion, shown separately for cusp and core galaxies.
Based on the \citet{LauerFaber+:2007a} sample we take $\rinf = 25(8)M
^{0.6}$ pc and $\rb = 106 \Mbheight^{0.4}$ (100 pc) and thus
$\rb/\rinf\simeq 4 \Mbheight^{-0.2} (13 \Mbheight^{-0.6})$ for cores
(cusps).  For cusp galaxies we see that the $v_w \gtrsim v_{\rm TI}$
criterion is generally more stringent, while for core galaxies the
$\zeta > \zeta_c$ criterion is more strigent for core galaxies, except
for high mass loss rates $\eta > 1$.

@@

Strikingly, we find that galactic nuclei that experience the same
average SFH as their host galaxies possess thermally stable flows
across all black hole masses.  An important caveat, however, is that
$\vw$ is generally only a few times higher than the stellar velocity
dispersion, i.e. $\zeta \sim \zeta_c$ (eq.~[\ref{eq:zetacrit}]).  This
implies that the true stagnation radius and the inflow rate could
be larger than our analytic estimates, potentially resulting in
thermal instability in a significant fraction of galaxies.  The bottom
panel of Figure~\ref{fig:vwSources} also shows $\zeta/\zeta_c$, where
$\zeta_c$ is estimated ...{\bf AG this sentence needs to be
  fixed. Power law fits different for cores and cusps. Power law fixed
to be 0.6 for influence radius in doing fits}.  In other words, the CNM of a
significant fraction of massive ellipticals may be thermally unstable,
not necessarily because they are receiving lower stellar feedback than
the average for their galaxy mass, but due to the high heating
required for stability given the structure of their stellar potential.

@@

% \begin{table*}
% \begin{threeparttable}
% \begin{minipage}{18cm}
% \caption{Comparison of CNM gas density at r=10$^{18}$ between our
%   model and existing constraints on the gas density on this scale for
%   the Milky Way from Chandra x-ray observations (first row) and from
%   modeling the radio light curve in the Swift J1644 transient (second row).}
% \begin{tabular}{lll}
%   {System} &  {Model} & {Comparison} \\
%   %\hline
%   Milky Way & 3 cm${-3}$ & 26 cm$^{-3}$\\
%   Swift J1644+57 & 5 cm$^{-3}$ & 0.8 cm$^{-3}$\\
%   %\hline
% \label{table:compare}  
% \end{tabular}
% % \begin{tablenotes}
% %   Model estimate has been corrected for the fact the analytic
% %   prescription underestimates gas density. This corresponds to model
% %   estimate for core galaxies.
% % \end{tablenotes}
% \end{minipage}
% \end{threeparttable}
% \end{table*}

@@

\begin{figure}
\includegraphics[width=\columnwidth]{mdot_sfr.pdf}
\caption{\label{fig:bh_growth} {\it Top panel:} Gas inflow rate
  $\dot{M}/\dot{M}_{\rm edd}$ as a function of black hole mass
  $M_{\bullet}$, shown for core ($\Gamma=0.1$, dot-dashed) and cusp
  galaxies ($\Gamma=0.8$, solid).  Black lines show the inflow rate
  calculated from equation (\ref{eq:eddr_analytic}) using the heating
  rate provided at $z = 0$ by the average star formation histories for
  each galaxy mass (Fig.~\ref{fig:vwSources}).  Teal lines show the
  maximum accretion set by Ia supernovae blow-out or heating,
  $\dot{M}_{\rm Ia}$ (\ref{eq:eddr_Ia}).  Red lines show the inflow
  rate obtained if Compton heating acts alone, $\dot{M}_{\rm C}$
  (eq.~[\ref{eq:MdotC}]).  Yellow lines show the maximum inflow
  rate for a thermally stable flow near the stagnation radius,
  $\dot{M}_{\rm TI}$ (eq.~[\ref{eq:Mdotmax}]).  Also shown are the
  inferred inflow rates of SgrA* from \citet{Quataert:2004a} (gray
  circles), for NGC3115 from \citet{ShcherbakovWong+:2014a} (gray
  diamonds), and for Swift J1644+57 from \citet{BergerZauderer+:2012a}
  (gray triangles).  {\it Bottom panel:} Growth times $t_{\rm grow}
  \equiv \Mbh/\dot{M}_{\bullet}$ in units of the Hubble time $\th$ for
  each of the accretion rates shown in the top panel, where
  $\dot{M}_{\bullet} = \alpha\dot{M}$, where $\alpha = 0.1$, to
  account for the fraction of inflowing mass lost to disk outflows on
  small scales.}
\end{figure}

@@

Shown also for comparison in Figure \ref{fig:bh_growth} (top panel) is
the inflow rate, $\dot{M}$, onto SgrA* calculated by
\citet{Quataert:2004a} and by \citet{Cuadra+2008}.  Also shown is the
range of $\dot{M}$ for the low-luminosity AGN NGC3115 derived through
detailed modeling by \citet{ShcherbakovWong+:2014a}, who find a range
of inflow rates depending on the assumed model for thermal
conductivity.  The inflow rate SgrA* estimated from
\citet{Quataert2004} exceeds our estimate of $\dot{M}$ due to the
average SFH for the same $M_{\bullet}$ by almost two orders of
magnitude. This is because the mass source term in
\citet{Quataert2004} exceeds ours by several orders of magnitude.
The star formation in the central parsec of the Milky Way cannot be
described as steady-state: feedback is instead dominated by the
stellar winds from the ring of young massive stars of mass $\sim
10^{4}M_{\odot}$ and estimated age $\sim 10$ Myr (e.g.,
\citealt{Schodel+07}).  Such a star formation history is better
described by our impulsive (starburst) scenario, for which the value
of $\eta \sim 100$ at $\tau_{\star} = 10^{7}$ yr
(Fig.~\ref{fig:vwImp}) is a factor of $\sim 100$ times higher than the
value of $\eta \sim 1$ predicted in the average SFH case
(Fig.~\ref{fig:vwSources}, bottom panel).

However, simulations by \citet{Cuadra+2006} find that
the inflow rate for Sgr A* is reduced by up to an order of magnitude
compared to the result in \citet{Quataert2004} when the discreteness
and disk geometry of stellar sources is taken into account.

@@

Our results also have implications for the environments encountered by
relativistic jets from TDEs.  For low-mass SMBHs with $M_{\bullet}
\lesssim 10^{7}M_{\odot}$, such as that responsible for powering the
transient {\it Swift} J1644+57 (\citealt{Bloom+11}), we predict gas
densities of a $n \sim 0.3-3$ cm$^{-3}$ on radial scales of 0.3 pc for
wind heating rates within the physically expected range $v_{\rm w}
\sim 500-1000$ km s$^{-1}$ (Fig.~\ref{fig:profiles}).  These values
are within the range estimated by modeling the radio afterglow of {\it
  Swift} J1644+57 (\citealt{Metzger+12},
\citealt{BergerZauderer+:2012a}), as shown in
Fig.~\ref{fig:bh_growth}.

@@

Our analysis of SMBH feedback has focused on Compton heating instead
of kinetic feedback, since the effects of radiative feedback are
relatively straightforward to calculate from the properties of the
accretion flow.  However, it is possible kinetic feedback from SMBH
outflows could play an equal or greater role in regulating accretion,
even in low luminosity AGN.

>>>>

\section{Model}
\label{sec:model}
\subsection{Galaxy models}
\label{sec:gal_model}
\citet{LauerFaber+:2007a} use Hubble Space Telescope WFPC2 imaging to
measure the radial surface brightness profiles for hundreds of nearby
early type galaxies. The measured profile is fit by a ``Nuker" law parameterization:
\begin{equation}
  I(\xi)=I_b 2^{(\beta-\Gamma)/\alpha} \xi^{-\Gamma} (1+\xi^\alpha)^{-(\beta-\Gamma)/\alpha}, \,\,\,\xi\equiv\frac{r}{r_b},
\end{equation}
i.e., a broken power law that transitions from an inner power law
slope, $\Gamma$, to an outer power law slope, $\beta$, at a break
radius, $\rb$.  If the stellar population is spherically symmetric,
then this corresponds to a 3D stellar density $\rhostar \propto
r^{-1-\Gamma}$ for $r \ll \rb$ and $\rhostar\propto r^{-1-\beta}$ for
$r \gg \rb$.  We can write the deprojected stellar density
approximately (formally, this is the $\alpha \rightarrow \infty$
limit) as
\begin{align}
\rho_\star = 
\begin{cases}
\rho_\star|_{\rinf} \left(r/\rinf\right)^{-1-\Gamma} & r \leq r_b\\
\rho_\star|_{r_b} \left(r/r_b\right)^{-1-\beta} & r > r_b,
\label{eq:rhostar}
\end{cases}
\end{align}
where $\rho_\star|_{\rinf}$ is the stellar density at the radius of
the black hole sphere of influence\footnote{This approximate
  expression agrees surprisingly well with the mean empirical scaling
  relation for $r_{\rm inf}(M_\bullet)$ calculated in Appendix C of
  \citet{Stone&Metzger15}, although we note that this relation has
  significant scatter. Also, we note that the  scaling is somewhat
  different for cores and cusps. Fixing the power law slope,
  $\rinf\simeq 25 (8) \Mbheight^{0.6}$ for core (cusp)
  galaxies. Unless otherwise noted we will use the single relation in
  equation~\eqref{eq:rsoi}},
\begin{equation}
  \rinf \simeq G \Mbh/\sigma_{\bullet}^2 \approx 14 M_{\bullet,8}^{0.6}\,{\rm pc},
\label{eq:rsoi}
\end{equation}
where $M_{\bullet,8} \equiv M_{\bullet}/10^{8}M_{\odot}$ and the
second equality in (\ref{eq:rsoi}) employs the $\Mbh-\sigma$
relationship of \citet{McConnellMa+:2011a},
 \begin{align}
M_{\bullet} \simeq 2\times 10^{8}\left(\frac{\sigma_{\bullet}}{200\,{\rm
      km\,s^{-1}}}\right)^{5.1}M_{\odot}.
\label{eq:Msigma}
\end{align}
This may be of questionable validity for low mass black holes (e.g.,
\citealt{Greene+2010, Kormendy+2013}). Also, several of the black hole
masses used in \citealt{McConnellMa+:2011a} were underestimated
\citep{Kormendy+2013}. However, our results are not overly sensitive
to the exact form of the $\Mbh-\sigma$ relationship that we use.

@@

We fix $\beta = 2$ but find that our results are not overly sensitive
to the properties of the gas flow on radial scales $\gtrsim r_b$.  The
presence of the break radius $r_{b}$ is, however, necessary to obtain
a converged steady state for some regions of our parameter space.  We
consider solutions calculated for up to four values of $\rb$: 50 pc,
100 pc, 200 pc, and 400 pc, motivated by the range of break radii from
the \citet{LauerFaber+:2007a} sample.\footnote{For the core galaxies
  the break radius follows the scaling relationship $\rb\sim 90 \,
  (\Mbheight)^{0.5}$ pc, with scatter of approximately one dex. Most
  the cusp galaxies have $\rb$ between 100 and 1000 pc, and lack a
  clear trend with $\Mbh$. The mean $r_b$ for cusp galaxies in the
  \citet{LauerFaber+:2007a} sample is $\sim 240$ pc} We neglect
values of $r_b$ which would give unphysically large bulges for a given
$\Mbh$.

@@

\begin{figure}
  \includegraphics[width=\columnwidth]{ti.pdf}
  \caption{Minimum effective wind heating parameter required for
    thermal stability as a function of SMBH mass.  Black lines show
    $v_{\rm TI}$ (eq.~[\ref{eq:cooling3}]), the heating rate required
    for $(\dot{q}_{\rm heat}/|\dot{q}_{\rm rad}|)_{r_{s}} > 10$ in the
    high-heating limit when the stagnation radius lies interior to the
    influence radius, for different values of the mass loss parameter
    $\eta$ as marked.  Blue lines show the minimum heating parameter
    required to have $\zeta > \zeta_c = (r_{b}/r_{\rm
      inf})^{0.5(1-\Gamma)}$ (eq.~[\ref{eq:zetacrit}]), separately for
    cusp (solid) and core (dashed) galaxies. Based on the
    \citet{LauerFaber+:2007a} sample we take $\rinf=25 (8) \Mbh^{0.6}
    pc$ and $\rb=90 \Mbheight^{0.5}$ pc (240 pc) for cores
    (cusps). For $\zeta < \zeta_c$ the stagnation radius moves from
    inside the influence radius, out to the stellar break radius $r_b$
    (Appendix \ref{app:be}); in practice renders the flow susceptible
    to thermal runaway, even if $v_{w} > v_{\rm TI}$.  
    % For cusp
%     galaxies the $v_w > v_{\rm TI}$ is usually the more stringent
%     criterion for stability, while for core galaxies the $\zeta >
%     \zeta_c$ is usually more relevant.
}
\label{fig:TI}
\end{figure}

@@

Figure \ref{fig:TI} shows $v_{\rm TI}(M_{\bullet})$ for different mass
input parameters $\eta$, as well as the [$\eta$-independent] $\zeta >
\zeta_c$ criterion, shown separately for cusp and core galaxies.
Based on the \citet{LauerFaber+:2007a} sample we take $\rinf = 25(8)M
^{0.6}$ pc and $\rb = 90 \Mbheight^{0.5}$ (240 pc) and thus
$\rb/\rinf\simeq 4 \Mbheight^{-0.1} (30 \Mbheight^{-0.6})$ for cores
(cusps).  For cusp galaxies we see that the $v_w \gtrsim v_{\rm TI}$
criterion is more stringent for $\eta \gsim 0.1$. The $\zeta >
\zeta_c$ is more important for core galaxies especially for $\Mbh
\gsim 10^8 \Msun$, where most core galaxies are found.

@@

Strikingly, we find that galactic nuclei that experience the same
average SFH as their host galaxies possess thermally stable flows
across all black hole masses.  An important caveat, however, is that
$\vw$ is generally only a few times higher than the stellar velocity
dispersion, i.e. $\zeta \sim \zeta_c$ (eq.~[\ref{eq:zetacrit}]).  This
implies that the true stagnation radius and the inflow rate could be
larger than our analytic estimates, potentially resulting in thermal
instability in a significant fraction of galaxies.  The bottom panel
of Figure~\ref{fig:vwSources} also shows $\zeta/\zeta_c$, where
$\zeta_c$ is estimated using fits for $r_b$, $\rinf$, and $\Gamma$
derived from the \citet{LauerFaber+:2007a} sample.  In other words,
the CNM of a significant fraction of massive ellipticals may be
thermally unstable, not necessarily because they are receiving lower
stellar feedback than the average for their galaxy mass, but due to
the high heating required for stability given the structure of their
stellar potential.

@@

@@

\begin{figure}
\includegraphics[width=\columnwidth]{mdot_sfr.pdf}
\caption{\label{fig:bh_growth} {\it Top panel:} Gas inflow rate
  $\dot{M}/\dot{M}_{\rm edd}$ as a function of black hole mass
  $M_{\bullet}$, shown for core ($\Gamma=0.1$, dot-dashed) and cusp
  galaxies ($\Gamma=0.8$, solid).  Black lines show the inflow rate
  calculated from equation (\ref{eq:eddr_analytic}) using the heating
  rate provided at $z = 0$ by the average star formation histories for
  each galaxy mass (Fig.~\ref{fig:vwSources}).  Teal lines show the
  maximum accretion set by Ia supernovae blow-out or heating,
  $\dot{M}_{\rm Ia}$ (\ref{eq:eddr_Ia}).  Red lines show the inflow
  rate obtained if Compton heating acts alone, $\dot{M}_{\rm C}$
  (eq.~[\ref{eq:MdotC}]).  Yellow lines show the maximum inflow rate
  for a thermally stable flow near the stagnation radius,
  $\dot{M}_{\rm TI}$ (eq.~[\ref{eq:Mdotmax}]).  Also shown are the
  inferred inflow rates of SgrA* from \citet{Quataert:2004a} and
  \citet{Cuadra+2008} (gray circles) and for NGC3115 from
  \citet{ShcherbakovWong+:2014a} (gray diamonds).  {\it Bottom panel:}
  Growth times $t_{\rm grow} \equiv \Mbh/\dot{M}_{\bullet}$ in units
  of the Hubble time $\th$ for each of the accretion rates shown in
  the top panel, where $\dot{M}_{\bullet} = \alpha\dot{M}$, where
  $\alpha = 0.1$, to account for the fraction of inflowing mass lost
  to disk outflows on small scales.}
\end{figure}

@@

Shown also for comparison in Figure \ref{fig:bh_growth} (top panel) is
the inflow rate, $\dot{M}$, onto SgrA* calculated by
\citet{Quataert:2004a} and by \citet{Cuadra+2008}.  Also shown is the
range of $\dot{M}$ for the low-luminosity AGN NGC3115 derived through
detailed modeling by \citet{ShcherbakovWong+:2014a}, who find a range
of inflow rates depending on the assumed model for thermal
conductivity.  The inflow rate SgrA* estimated from
\citet{Quataert:2004a} exceeds our estimate of $\dot{M}$ due to the
average SFH for the same $M_{\bullet}$ by almost two orders of
magnitude. This is because the mass source term in
\citet{Quataert:2004a} exceeds ours by several orders of magnitude.
The star formation in the central parsec of the Milky Way cannot be
described as steady-state: feedback is instead dominated by the
stellar winds from the ring of young massive stars of mass $\sim
10^{4}M_{\odot}$ and estimated age $\sim 10$ Myr (e.g.,
\citealt{Schodel+07}).  Such a star formation history is better
described by our impulsive (starburst) scenario, for which the value
of $\eta \sim 100$ at $\tau_{\star} = 10^{7}$ yr
(Fig.~\ref{fig:vwImp}) is a factor of $\sim 100$ times higher than the
value of $\eta \sim 1$ predicted in the average SFH case
(Fig.~\ref{fig:vwSources}, bottom panel).

However, simulations by \citet{Cuadra+2006} find that
the inflow rate for Sgr A* is reduced by up to an order of magnitude
compared to the result in \citet{Quataert:2004a} when the discreteness
and disk geometry of stellar sources is taken into account.

@@

Our results also have implications for the environments encountered by
relativistic jets from TDEs.  For low-mass SMBHs with $M_{\bullet}
\lesssim 10^{7}M_{\odot}$, such as that responsible for powering the
transient {\it Swift} J1644+57 (\citealt{Bloom+11}), we predict gas
densities of a $n \sim $few cm$^{-3}$ on radial scales of 0.3 pc for
wind heating rates within the physically expected range $v_{\rm w}
\sim 500-1000$ km s$^{-1}$ and $\eta=0.4$ {\bf AG might be good to
  cite figure 2 here but this has eta=0.2}.  These values are within
the range estimated by modeling the radio afterglow of {\it Swift}
J1644+57 (\citealt{Metzger+12},\citealt{BergerZauderer+:2012a}). However,
\citet{Mimica+2015} find a much higher density of $60 cm^{-3}$ at 0.3 pc.

@@

Our analysis of SMBH feedback has focused on Compton heating instead
of kinetic feedback, since the effects of radiative feedback are
relatively straightforward to calculate from the properties of the
accretion flow.  However, it is possible kinetic feedback from SMBH
outflows could play an equal or greater role in regulating accretion,
even in low luminosity AG.

### 2015-07-02 windheat.tex
sclaing relations and caveats

\begin{equation}
\dot{\mathcal{M}}=10^{-5.4} \left(\frac{t}{4\times 10^6 {\rm
       yr}}\right)^{-3} 4 {\, \rm Myr} \leq t \leq 40 {\, \rm Myr}. 
\end{equation}

and 0 for all later times. We neglect the mass injection due to
stellar winds for times $t\lsim 4$ Myr. $\dot{m}_{\rm OB}=f_{8}
\dot{\mathcal{M}}/\bar{m}_{\star}$, where $f_{8} =2.6 \times 10^{-3}$
is the fraction of the stellar mass with $M_{\star} > 8M_{\odot}$ and
$m_{\star}=0.35 \Msun$ is the mean stellar mass for our assumed
Salpeter IMF, $\mu\sim M_\star^{-2.35}$.

@@

and 0 for earlier times. By truncating $dot{m}_{\rm TO}$ at 40 Myr, we
are ignoring the non-steady mass injection by type II supernovae.

@@

The rate of energy injection due to winds from massive
stars, $\dot{e}_{\rm OB}(t) = f_{8}\dot{\mathcal{E}}/\bar{m}_*$, where
$f_{8} =2.6 \times 10^{-3}$ is the fraction of the stellar mass with
$M_{\star} > 8M_{\odot}$ for our assumed Salpeter IMF.  Here
$\dot{\mathcal{E}} (t)$ is the energy injection rate per massive star,
which we estimate as
\begin{align}
\dot{\mathcal{E}} (t)=  1.3 \times 10^{36} {\rm erg\,s^{-1}}
\begin{cases}
  1 & t<4 \times 10^6 {\rm yr}\\
  \left(\frac{t}{4 \times  10^6\,{\rm yr}}\right)^{-3.73} &  4\times
  10^{6} {\rm yr}  \le t \le 10^{8} {\rm yr}
  \times 10^6 {\rm yr},\\
  0 & t > 10^8 {\rm yr}
\end{cases}
\label{eq:voss}
\end{align}
based on the results of \citet{VossDiehl+:2009a} (their Fig.~7, top
panel), who use a population synthesis code to simulate the mass and
energy injection into the ISM from an OB association. {\bf AG--should
  just truncate the energy injection at $10^7$ years. Explain why
  turn-off energy injection is negligible in lower mass stars.}

>>>>

\begin{equation}
\dot{\mathcal{M}}=
\begin{cases}
10^{-6.5} \Msun {\rm yr^{-1}} & t < 4 Myr\\
10^{-5.4} \Msun {\rm yr^{-1}} \left(\frac{t}{4\times 10^6 {\rm
       yr}}\right)^{-3} & 4 {\, \rm Myr} \leq t \leq 40 {\, \rm Myr}. \\
0 & t>40 {\rm Myr}
\end{cases}
\end{equation}

We neglect the mass injection due to stellar winds for times $t\lsim
4$ Myr. $\dot{m}_{\rm OB}=f_{8} \dot{\mathcal{M}}/\bar{m}_{\star}$,
where $f_{8} =2.6 \times 10^{-3}$ is the fraction of the stellar mass
with $M_{\star} > 8M_{\odot}$ and $m_{\star}=0.35 \Msun$ is the mean
stellar mass for our assumed Salpeter IMF, $\mu\sim M_\star^{-2.35}$.

@@

and 0 for earlier times. By truncating $\dot{m}_{\rm TO}$ at 40 Myr, we
are ignoring the non-steady mass injection by type II supernovae.

@@

The rate of energy injection due to winds from massive
stars, $\dot{e}_{\rm OB}(t) = f_{8}\dot{\mathcal{E}}/\bar{m}_*$, where
$f_{8} =2.6 \times 10^{-3}$ is the fraction of the stellar mass with
$M_{\star} > 8M_{\odot}$ for our assumed Salpeter IMF.  Here
$\dot{\mathcal{E}} (t)$ is the energy injection rate per massive star,
which we estimate as
\begin{align}
\dot{\mathcal{E}} (t)=  1.3 \times 10^{36} {\rm erg\,s^{-1}}
\begin{cases}
  1 & t<4 \times 10^6 {\rm yr}\\
  \left(\frac{t}{4 \times  10^6\,{\rm yr}}\right)^{-3.73} &  4\times
  10^{6} {\rm yr}  \le t \le 10^{8} {\rm yr},\\
  0 & t > 10^8 {\rm yr}
\end{cases}
\label{eq:voss}
\end{align}
based on the results of \citet{VossDiehl+:2009a} (their Fig.~7, top
panel), who use a population synthesis code to simulate the mass and
energy injection into the ISM from an OB association. {\bf AG--should
  just truncate the energy injection at $10^7$ years. Explain why
  turn-off energy injection is negligible in lower mass stars.}
